\stdpopsim: a community toolkit for population genetic simulation

### 2019-12-06 popSimManu.tex

\subsection*{Workflow for analysis of simulated data}
Snakemake pipeline for 1 pop models
Snakemake pipeline for 2 pop models

 In general, used default
settings and did not attempt to optimize \dadi or \fastsimcoal runs Used Snakemake to
manage pipeline for two population analysis Results from two population pipeline
demonstrate that the different methods tested were generally able to recover the
simulated parameters

>>>>

\subsection*{Workflow for analysis of simulated data}
Snakemake pipeline for 1 pop models

\subsection*{Workflow for two population analysis}
Our goal for the two population analysis pipeline was to explore two population inference
on some of the multi-population demographic models implemented in \stdpopsim.
The models we examined include the human Gutenkunst model (\cite{gutenkunst2009inferring}) as
well as the \emph{Drosophila} Li & Stephan model (\cite{li2006inferring}). The human
model is a three population model (Africa, Europe, and Asia) including post-divergence
migration and exponential growth (Figure \ref{fig:IM_popn_human}C), whereas the
\emph{Drosophila} model is a two population model (Africa and Europe) with no post-divergence
migration and constant population sizes (Figure Sx).

To conduct inference on these models, we implemented three commonly-used approaches:
\dadi (\cite{gutenkunst2009inferring}), \fastsimcoal2 (\cite{excoffier2013robust}),
and \smcpp (\cite{terhorst2017robust}). Our aim was not to exhaustively
test the performance of these methods, but rather to explore their behavior
in two different inference scenarios and demonstrate the potential for this resource to
eventually be used for more in-depth benchmarking. As a result, these methods were used
generally with default settings and we did not attempt to optimize their performance or fit
parameter-rich demographic models. As with the single population inference pipeline,
we developed a \texttt{snakemake} (\cite{koster2012snakemake}) workflow to manage
our inference pipeline and maintain reproducibility.

For both \dadi and \fastsimcoal2, our general approach was to fit a two population
isolation with migration (IM) model with constant population sizes to the simulated
data from the human and \emph{Drosophila} models. This IM model contains six parameters:
the ancestral population size, the size of population 1 after the split, the size of
population 2 after the split, the divergence time, and two migration rate parameters.
Importantly, this meant that for both the human and \emph{Drosophila} inference, the
fitted model did not match the simulated model (Figure \ref{fig:IM_popn_human}C and Sx).
In the human case, we therefore performed inference solely on the Africa
and Europe populations, meaning that the Asia population functioned as a ghost
population that was ignored by our inference. Our motivation for fitting this simple
IM model was to mimic the typical approach of two population inference on empirical
data, where the user is not aware of the 'true' underlying demography and the inference
model is often misspecified. To ground-truth our inference approach, we also conducted
inference on a generic IM model that was identical to the model used for inference.

To perform inference, we began by running simulations under each model, running
three replicates to inspect for inter-replicate variability. For humans, we ran
whole genome simulations (chromosomes 1-22) under the Gutenkunst model (\cite{gutenkunst2009inferring}),
using the HapMap II recombination map (\cite{international2007second}) and taking 20
samples each from the Europe and Africa populations. For \emph{Drosophila},
we simulated 3 Mb of chromosome 2R under the Li & Stephan model (\cite{li2006inferring}),
using the map inferred by \cite{comeron2012many} and taking 50 samples from each of
the Africa and Europe populations. We tried to run whole genome simulations under
this model but after XXXX amount of time... (JRA fill in details here). For the generic
IM simulations, we used the human genome (chromosomes 1-22) and HapMap II recombination
map (\cite{international2007second}) and took 20 samples per population.

Following simulation, we outputted tree sequences and masked low recombination
regions using the same approach described for the single population pipeline above. We
converted tree sequences into a two dimensional site frequency spectrum for all
chromosomes in the appropriate format for \dadi and \fastsimcoal2. For each simulation
replicate, we performed 10 runs of \dadi and \fastsimcoal2 and checked for convergence.
Detailed settings for \dadi and \fastsimcoal2 can be found in the Snakefile
on the popgensims/analysis GitHub repository (https://github.com/popgensims/analysis). The parameter
estimates from the top run from each simulation replicate are shown in Figures \ref{fig:IM_popn_human}C and Sx.

For \smcpp , we converted the tree sequences into vcf format and performed inference
with default settings. Importantly, \smcpp assumes no migration post-divergence, so
here again the inference model does not exactly match the simulated model for the
human model that does include migration. However, because \smcpp allows
for continuous population size changes, it is better equipped to capture many of the
more complex aspects of the simulated demographic models (e.g., exponential growth).

To visualize our results, we plotted the inferred population size trajectories
for each simulation replicate along side the true (census) population sizes
(Figures \ref{fig:IM_popn_human}C Sx). Here, unlike the single population pipeline,
 we focus on census size rather than the inverse coalescence rate as the 'true' population size.
